---
title: "Vorticity"
author: "Sirish Shrestha"
date: "August 25, 2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Echo Vorticity Analysis

#### Read the data for vorticity analysis
```{r cars}

setwd("C:/Users/siris/Documents/rstudio-projects/vorticity")

## read all echo data
echo <- read.csv("data.csv", header = TRUE, sep = ",", stringsAsFactors = TRUE)

ncol(echo)

#summary(echo)
#str(echo)
## There are several missing data. We only select the ones we care about - the vortex data.
vorticity.data <- echo[,c(45:52)]


#
plot(vorticity.data)



mar <- par()$mar
par(mar=mar+c(0,5,0,0))

# seems like the variables are in different scales
barplot(sapply(vorticity.data, var), horiz=T, las=1, cex.names=0.8)

# in log scale
barplot(sapply(vorticity.data, var), horiz=T, las=1, cex.names=0.8, log='x')
par(mar=mar)


pc.unscaled <- princomp(vorticity.data)
summary(pc.unscaled)
loadings(pc.unscaled)
plot(pc.unscaled)


# Scale
data2 <- data.frame(scale(vorticity.data))
# Verify variance is uniform
plot(sapply(data2, var))


```





## Calculating PCA

```{r}

# Proceed with principal components
pc <- princomp(data2)
plot(pc)

plot(pc, type='l')
summary(pc) # 4 components is both 'elbow' and explains >85% variance
loadings(pc)


# Get principal component vectors using prcomp instead of princomp
pc <- prcomp(data2)

# First for principal components
comp <- data.frame(pc$x[,1:4])
# Plot
plot(comp, pch=16, col=rgb(0,0,0,0.5))

```


```{r}
library(rgl)
# Multi 3D plot
plot3d(comp$PC1, comp$PC2, comp$PC3)
plot3d(comp$PC1, comp$PC3, comp$PC4)


```

###Hierarchical Clustering packages


#### Load packages
```{r, include=FALSE}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(cluster))
suppressPackageStartupMessages(library(factoextra))
suppressPackageStartupMessages(library(dendextend))

library(tidyverse)  # data manipulation
library(cluster)    # clustering algorithms
library(factoextra) # clustering visualization
library(dendextend) # for comparing two dendrograms

```


```{r}

# Dissimilarity matrix
d <- dist(comp, method = "euclidean")

# Hierarchical clustering using Complete Linkage
hc1 <- hclust(d, method = "complete" )

# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)



```



## For Hierarchical Clustering

#### using Agnes to check the cluster structure

```{r, echo=TRUE}
library(cluster)
library(purrr)

# methods to assess
m <- c( "average", "single", "complete", "ward")
names(m) <- c( "average", "single", "complete", "ward")

# function to compute coefficient
ac <- function(x) {
  agnes(comp, method = x)$ac
}

map_dbl(m, ac)
##  average    single  complete      ward 
##0.7656333 0.6595636 0.8810995 0.9480419 


```

### we select Ward's method since it's the strongest structure

```{r}

library(dplyr)
require(graphics)

# Ward's method
hc5 <- hclust(dist(comp, method = "euclidean"), method = "ward.D2" )
plot(hc5, cex = 0.6, hang = -1)

# Cut tree into 4 groups
k <- 4
sub_grp <- cutree(hc5, k = k)

echo.cluster <- echo %>%
mutate(cluster = sub_grp)

# Cluster centroids

# function to find medoid in cluster i
#echo.cluster.centroid = function(i, dat, sub_grp) {
#    ind = (sub_grp == i)
#    colMeans(dat[ind,])
#}

#sapply(unique(sub_grp), echo.cluster.centroid, echo.cluster, sub_grp)

#apply(echo.cluster, 2, function (x) tapply (x, sub_grp, mean, na.rm=TRUE), na.rm=TRUE)


#echo.cluster$cluster <- as.factor(echo.cluster$cluster)


plot(hc5, cex=0.6)
rect.hclust(hc5, k = k, border = 2:5)

fviz_cluster(list(data = vorticity.data, cluster = sub_grp))


```

#Analyzing 3rd cluster
```{r}

cluster3 <- echo.cluster[echo.cluster$cluster == 3, ]
cluster2 <- echo.cluster[echo.cluster$cluster == 2, ]



```
#PCA plot with cluster

```{r}

library(RColorBrewer)

palette(alpha(brewer.pal(9, 'Set1'), 0.5))
plot(comp, col=sub_grp, pch=16)



# 3D plot
plot3d(comp$PC1, comp$PC2, comp$PC3, col=sub_grp)
plot3d(comp$PC1, comp$PC3, comp$PC4, col=sub_grp)


```



```{r}


# Cluster sizes
sort(table(sub_grp))
clust <- names(sort(table(sub_grp)))

# First cluster
#row.names(echo.cluster[sub_grp==clust[1],])
# Second Cluster
#row.names(echo.cluster[sub_grp==clust[2],])
# Third Cluster
#row.names(echo.cluster[sub_grp==clust[3],])

for(i in 1:k){
  cat(paste('# Cluster', i, "\n"))
  r <- row.names(echo.cluster[sub_grp==clust[i],])
  cat(r, "\n\n")
}

```


# compare vars in 

```{r}

echo.cluster[echo.cluster$cluster==unique(sub_grp)[3],c(10,24)]

echo.cluster[echo.cluster$cluster==unique(sub_grp)[4],c(10,24)]



```

# Analysis of demographics with cluster
# Chi square Test
      
```{r}

echo.chisq.fn <- function(x){
  
  for(i in 1:length(x)){
    
    chi.var.name <- colnames(echo.cluster[x[i]])
    
    echo.table[[chi.var.name]] <- table(cluster = echo.cluster$cluster, Var = echo.cluster[[chi.var.name]])
    
    cat("Var name: ", chi.var.name, "\n")
    cat("------------------------------\n\n")
    
    cat("2-way Table\n")
    cat("============\n")
    print(echo.table[[chi.var.name]])
    cat("\n\n\n")
    
    cat("Chi-Square test\n")
    cat("================\n\n")
    echo.chisq.test <- chisq.test(echo.table[[chi.var.name]])
    
    print(echo.chisq.test)
    
    cat("\n\n")
    boxplot(echo.cluster[[chi.var.name]] ~ echo.cluster$cluster, 
            xlab = "Cluster", 
            ylab = chi.var.name, 
            main = paste("Cluster by ", chi.var.name))
    
    cat("\n\n=======================================================================\n\n")
  }
  
  
} 


echo.table <- list()
chisq.cols <- c(4, 9, 24, 61:64)
echo.chisq.fn(chisq.cols)



```



```{r}

#age range
echo.cluster$age.range <- cut(echo.cluster$Age, breaks=c(0,40, 60, 80, 200), right = FALSE, labels = c("1", "2", "3", "4"))


# Age range by cluster
# 1 : 0-39
# 2:  40 - 59
# 3:  60 - 79
# 4:  80 +

age.table <- table(echo.cluster$cluster, echo.cluster$age.range)

chisq.test(age.table)


```



# ANOVA of all continuous variables
```{r}


echo.aov.fn <- function(x){
    
    for(i in 1:length(x)){
      
      var.name <- colnames(echo.cluster[x[i]])
      echo.aov[[var.name]] <- aov(echo.cluster[[var.name]] ~ as.factor(echo.cluster$cluster))
      
      # Show the variable name
      cat("==============================================================================\n")
      cat(var.name,"\n")
      
      
      # Print the p-value
      cat(paste("P-value: ", summary(echo.aov[[var.name]])[[1]][["Pr(>F)"]][[1]]))
      
      
        if(summary(echo.aov[[var.name]])[[1]][["Pr(>F)"]][[1]] <= 0.05){
          
          cat("\n\nANOVA\n")
          cat("------\n")
          cat('var.name: ', var.name, '\n')
          cat("=============================\n\n")
          
          print(echo.aov[[var.name]])
          
          cat("\n\n")
          # Perform Tukey HSD Test
          tukey <- TukeyHSD(echo.aov[[var.name]])
          print(tukey)
          
        }
        else{
          cat("Not Significant\n\n")
        }
      
      
      cat("==============================================================================\n\n\n\n") 
    
     }
  
    return("Done")
}
echo.aov <- list()
columns <- c(3, 5:8,11:23,25:35)
echo.aov.fn(columns)



# Significant
# EchoLA, EchoDd, EchoDs, EchoEDV, EchoESV, EchoEF, TimeTMFDecT, EchSepTDIe, EchoAveTDIe, EchoLVM, echolvmass

```















