---
title: "Vorticity - non PCA"
author: "Sirish Shrestha"
date: "August 25, 2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Echo Vorticity Analysis


#### Load packages
```{r, include=FALSE}

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(cluster))
suppressPackageStartupMessages(library(factoextra))
suppressPackageStartupMessages(library(dendextend))
suppressPackageStartupMessages(library(rcompanion))
suppressPackageStartupMessages(library(purrr))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(graphics))
suppressPackageStartupMessages(library(RColorBrewer))
suppressPackageStartupMessages(library(fifer))
suppressPackageStartupMessages(library(fpc))

library(tidyverse)  # data manipulation
library(cluster)    # clustering algorithms
library(factoextra) # clustering visualization
library(dendextend) # for comparing two dendrograms
library(rcompanion)
library(purrr)
library(dplyr)
library(graphics)
library(RColorBrewer)
library(fifer)
library(fpc)

```

#### Read the data for vorticity analysis
```{r vorticity}



setwd("C:/Users/siris/Documents/rstudio-projects/vorticity")

## read all echo data
echo <- read.csv("data.csv", header = TRUE, sep = ",", stringsAsFactors = TRUE)

#echo$Gender <- factor(echo$Gender)
#echo$NYHA <- factor(echo$NYHA)
#echo$Etiology <- factor(echo$Etiology)

names(echo)[1] <- "ID"

echo$Gender <- factor(echo$Gender)
echo$NYHA <- factor(echo$NYHA)
echo$Etiology <- factor(echo$Etiology)
echo$DiastCount <- factor(echo$DiastCount)
echo$DiastDysfunASE <- factor(echo$DiastDysfunASE)
echo$DiastDysfunction3of4 <- factor(echo$DiastDysfunction3of4)
echo$Diastolic4count <- factor(echo$Diastolic4count)
echo$LAVI35 <- factor(echo$LAVI35)
echo$SeptTDIe7 <- factor(echo$SeptTDIe7)
echo$LatTDIe10 <- factor(echo$LatTDIe10)
echo$EtoTDIe14 <- factor(echo$EtoTDIe14)
echo$TRVmax <- factor(echo$TRVmax)
echo$Tracking <- factor(echo$Tracking)
echo$hr <- 60000/echo$TimePIVRRms

dim(echo)

str(echo)

summary(echo)
sapply(echo, function(x) round(mean(is.na(x)), 2))


#remove the diastolic data and ID for heatmap
echo.matrix <- data.matrix(echo[,-c(1,61:64)])
echo.matrix <- scale(echo.matrix)

summary(echo.matrix)

heatmap(echo.matrix, Rowv = NA, Colv = NA)

## There are several missing data. We only select the ones we care about - the vortex data.
vorticity.data <- echo.matrix[,c(44:51)]

cor(vorticity.data)

summary(vorticity.data)

#scaled voriticity
heatmap(vorticity.data)


#
plot(data.frame(vorticity.data))


# seems like the variables are in different scales
#barplot(sapply(data.frame(vorticity.data), var), horiz=T, las=1, cex.names=0.8)

# in log scale
#barplot(sapply(data.frame(vorticity.data), var), horiz=T, las=1, cex.names=0.8, log='x')
#par(mar=mar)



```


###Hierarchical Clustering packages

```{r}

# Dissimilarity matrix
d <- dist(vorticity.data, method = "euclidean")

# Hierarchical clustering using Complete Linkage
hc1 <- hclust(d, method = "complete" )

# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)



```



## For Hierarchical Clustering

#### using Agnes to check the cluster structure

```{r, echo=TRUE}


# methods to assess
m <- c( "average", "single", "complete", "ward")
names(m) <- c( "average", "single", "complete", "ward")

# function to compute coefficient
ac <- function(x) {
  agnes(vorticity.data, method = x)$ac
}

map_dbl(m, ac)
##  average    single  complete      ward 
##0.7656333 0.6595636 0.8810995 0.9480419 


```

### we select Ward's method since it's the strongest structure

```{r}

vorticity.data.frame <- data.frame(vorticity.data)

# distance
d <- dist(vorticity.data.frame, method = "euclidean")

# Ward's method
fit <- hclust(d, method = "ward.D2" )

#single
# fit <- hclust(d, method = "complete" )
plot(fit, cex = 0.6, hang = -1)




#Silhouette analysis for determining the number of clusters

asw <- numeric(20)
for (k in 2:20)
  asw[[k]] <- pam(vorticity.data.frame, k) $ silinfo $ avg.width
k.best <- which.max(asw)

plot(pam(d,k.best), color = c(1,2,3))

# Cut tree into 4 groups
k <- k.best
sub_grp <- cutree(fit, k = k)

#echo.cluster <- echo %>%
#mutate(cluster = sub_grp)

vorticity.data.cluster <- echo %>% mutate(cluster = sub_grp)

#age range


# Age range by cluster
# 1 : 0-39
# 2:  40 - 59
# 3:  60 - 79
# 4:  80 +

vorticity.data.cluster$age.range <- cut(vorticity.data.cluster$Age, breaks=c(0,40, 60, 80, 200), right = FALSE, labels = c("<40", "40-60", "60-80", "80+"))



plot(fit, cex=0.6)
rect.hclust(fit, k = k, border = 2:5)

fviz_cluster(list(data = vorticity.data, cluster = sub_grp))


```



```{r}


palette(alpha(brewer.pal(9, 'Set1'), 0.5))
plot(vorticity.data, col=sub_grp, pch=16)



```



```{r}


# Cluster sizes
sapply(split(vorticity.data.cluster, sub_grp), nrow)


# on scaled data
vorticity.cluster <- split(data.frame(vorticity.data), sub_grp)

(vorticity.cluster.colmeans <- sapply(vorticity.cluster, colMeans))


# on unsacled data

vorticity.cluster.unscaled <- split(data.frame(vorticity.data.cluster[,c(45:52, 65)]), sub_grp)
(voriticy.cluster.unscaled.colmeans <- (sapply(vorticity.cluster.unscaled, colMeans)))

clust <- names(sort(table(sub_grp)))

# First cluster
#row.names(echo.cluster[sub_grp==clust[1],])
# Second Cluster
#row.names(echo.cluster[sub_grp==clust[2],])
# Third Cluster
#row.names(echo.cluster[sub_grp==clust[3],])

for(i in 1:k){
  cat(paste('# Cluster', i, "\n"))
  r <- row.names(vorticity.data.cluster[sub_grp==clust[i],])
  cat(r, "\n\n")
}

```


# compare vars in 

```{r}

a <- vorticity.data.cluster[vorticity.data.cluster$cluster==unique(sub_grp)[1],c(10,24)]
b <- vorticity.data.cluster[vorticity.data.cluster$cluster==unique(sub_grp)[2],c(10,24)]
c <- vorticity.data.cluster[vorticity.data.cluster$cluster==unique(sub_grp)[3],c(10,24)]
d <- vorticity.data.cluster[vorticity.data.cluster$cluster==unique(sub_grp)[4],c(10,24)]



```

## checking 3rd cluster
```{r}

cluster3 <- vorticity.data.cluster[vorticity.data.cluster$cluster == 3,]

# cluster3$Gender <- factor(cluster3$Gender)
# cluster3$NYHA <- factor(cluster3$NYHA)
# cluster3$Etiology <- factor(cluster3$Etiology)

xtabs(~ age.range + Gender, data = cluster3)

summary(cluster3)

```


# Analysis of demographics with cluster
# Chi square Test
      
```{r}
echo.table <- list()

echo.chisq.fn <- function(x){
  
  
  for(i in 1:length(x)){
  
    chi.var.name <- colnames(vorticity.data.cluster[x[i]])
  
    echo.table[[chi.var.name]] <- xtabs(~ cluster + vorticity.data.cluster[[chi.var.name]], data = vorticity.data.cluster)
    
    cat("Var name: ", chi.var.name, "\n")
    cat("------------------------------\n\n")
    
    cat("2-way Table\n")
    cat("============\n")
    print(echo.table[[chi.var.name]])
    cat("\n\n\n")
    
    cat("Chi-Square test\n")
    cat("================\n\n")
    cat(chi.var.name)
    echo.chisq.test <- chisq.test(echo.table[[chi.var.name]])
    
    chisq.p.value <- echo.chisq.test$p.value
    
    if(chisq.p.value <= 0.05)
    {
    
      print(echo.chisq.test)
      
      cat("\n\n")
      
      mosaicplot(echo.table[[chi.var.name]], xlab="Cluster", ylab=chi.var.name, main = paste("Cluster-", chi.var.name, " mosaic plot"))
      
      boxplot(as.numeric(as.character(vorticity.data.cluster[[chi.var.name]])) ~ vorticity.data.cluster$cluster, 
              xlab = "Cluster", 
              ylab = chi.var.name, 
              main = paste("Cluster by ", chi.var.name))
      
      cat("\n\n=======================================================================\n\n")
    }
  }
  
  
} 



chisq.cols <- c(4, 9, 10, 24, 44, 56:64,66)
echo.chisq.fn(chisq.cols)



chisq.post.hoc(xtabs(~ cluster + NYHA, vorticity.data.cluster), control = "bonferroni")

chisq.post.hoc(xtabs(~ cluster + AbnormalEF50, vorticity.data.cluster), control = "bonferroni")


```



```{r}


cat("Age range by cluster
# 1 : 0-39
# 2:  40 - 59
# 3:  60 - 79
# 4:  80 +")

age.table <- table(cluster = vorticity.data.cluster$cluster, age.range = vorticity.data.cluster$age.range)
age.table

chisq.test(age.table)


```



```{r}
# 
# echo.aov.fn <- function(x){
#     
#     for(i in 1:length(x)){
#       
#       var.name <- colnames(vorticity.data.cluster[x[i]])
#       echo.aov[[var.name]] <- aov(vorticity.data.cluster[[var.name]] ~ as.factor(vorticity.data.cluster$cluster))
#       
#       # Show the variable name
#       cat("==============================================================================\n")
#       cat(var.name,"\n")
#       
#       
#       # Print the p-value
#       cat(paste("P-value: ", summary(echo.aov[[var.name]])[[1]][["Pr(>F)"]][[1]]))
#       
#       
#         if(summary(echo.aov[[var.name]])[[1]][["Pr(>F)"]][[1]] <= 0.05){
#           
#           cat("\n\nANOVA\n")
#           cat("------\n")
#           cat('var.name: ', var.name, '\n')
#           cat("=============================\n\n")
#           
#           print(summary(echo.aov[[var.name]]))
#           
#           cat("\n\n")
#           # Perform Tukey HSD Test
#           tukey <- TukeyHSD(echo.aov[[var.name]])
#           print(tukey)
#           
#         }
#         else{
#           cat("  Not Significant\n\n")
#         }
#       
#       
#       cat("==============================================================================\n\n\n\n") 
#     
#      }
#   
#     return("Done")
# }
# echo.aov <- list()
# columns <- c(3, 5:8,11:23,25:35)
# echo.aov.fn(columns)
# 
# 
# 
# 
# 
# 
# 
# 

```


# ANOVA of all continuous variables
```{r}
library(FSA)

echo.aov.fn <- function(x){
    
    for(i in 1:length(x)){
      
      var.name <- colnames(vorticity.data.cluster[x[i]])

      
      k.test <- kruskal.test(vorticity.data.cluster[[var.name]] ~ as.factor(vorticity.data.cluster$cluster))
      
      # Show the variable name
      cat("==============================================================================\n")
      cat(var.name,"\n")
      
      p.value <- k.test$p.value
      
      boxplot(vorticity.data.cluster[[var.name]] ~ factor(cluster), data=vorticity.data.cluster, xlab="Cluster", ylab=var.name, main=paste(var.name, " by cluster"))
      
      # Print the p-value
      cat(paste("P-value: ", p.value))
      
      
        if(p.value <= 0.05){
          
          cat("\n\n Kruskal-Wallis ran sum test \n")
          cat("--------------------------\n")
          cat('var.name: ', var.name, '\n')
          cat("=============================\n\n")
          
          print(k.test)
          
          cat("\n\n")

          dunn <- dunnTest(vorticity.data.cluster[[var.name]] ~ factor(cluster), data = vorticity.data.cluster, method ="bonferroni", na.action(na.omit(NA)))
          
          print(dunn)
          
        }
        else{
          cat("  Not Significant\n\n")
        }
      
      
      cat("==============================================================================\n\n\n\n") 
    
     }
  
    return("Done")
}


columns <- c(3, 5:8,11:23,25:36,53:55,65)
echo.aov.fn(columns)



# Significant
# EchoLA, EchoDd, EchoDs, EchoEDV, EchoESV, EchoEF, TimeTMFDecT, EchSepTDIe, EchoAveTDIe, EchoLVM, echolvmass

```















